---
title: Research tools
subtitle: robjhyndman/research\_tools
author: Rob J Hyndman
date: 2024-12-30
toc: true
titlegraphic: bg-12.png
titlecolor: Orange
format:
  presentation-beamer:
    pdf-engine: pdflatex
include-in-header:
  text: |
    \usepackage{fontawesome,pifont}
execute:
  echo: true
---

# Citing

## Managing references

\placefig{8.7}{2.6}{width=6.1cm}{Zotero}

\begin{textblock}{7.}(.4,0.95)
\begin{alertblock}{Zotero\phantom{y}}\fontsize{12}{12}\sf
\begin{itemize}\itemsep=1.mm
\item[\textcolor{white}{\ding{229}}] Free and on all operating systems
\item[\textcolor{white}{\ding{229}}] Web-version and local version synced
\item[\textcolor{white}{\ding{229}}] Browser extension for adding papers/books
\item[\textcolor{white}{\ding{229}}] Attach notes and annotations to papers.
\item[\textcolor{white}{\ding{229}}] Works with Word, LibreOffice or LaTeX.
\item[\textcolor{white}{\ding{229}}] Generate bibliography automatically
\item[\textcolor{white}{\ding{229}}] Handles all formatting for you.
\end{itemize}
\end{alertblock}
\end{textblock}

\begin{textblock}{7.7}(7.9,6.5)\fontsize{12}{13}\sf
\begin{block}{To install:}
\begin{itemize}
\item Set up account at \url{www.zotero.org}
\item Download from \url{www.zotero.org}
\end{itemize}
\end{block}
\end{textblock}\vspace*{10cm}

## Managing references

\placefig{9.2}{1.25}{width=5.3cm}{Mendeley}

\begin{textblock}{7.}(.4,0.95)
\begin{alertblock}{Mendeley}\fontsize{12}{12}\sf
\begin{itemize}\itemsep=1.mm
\item[\textcolor{white}{\ding{229}}] Free and on all operating systems
\item[\textcolor{white}{\ding{229}}] Web-version and local version synced
\item[\textcolor{white}{\ding{229}}] Browser extension for adding papers/books
\item[\textcolor{white}{\ding{229}}] Attach notes and annotations to papers.
\item[\textcolor{white}{\ding{229}}] Works with Word, LibreOffice or LaTeX.
\item[\textcolor{white}{\ding{229}}] Generate bibliography automatically
\item[\textcolor{white}{\ding{229}}] Handles all formatting for you.
\end{itemize}
\end{alertblock}
\end{textblock}

\begin{textblock}{7.7}(7.9,6.5)\fontsize{12}{13}\sf
\begin{block}{To install:}
\begin{itemize}
\item Set up account at \url{mendeley.com}
\item Download from \url{mendeley.com}
\end{itemize}
\end{block}
\end{textblock}\vspace*{10cm}

## Managing references

\placefig{8.7}{2}{width=6.1cm}{Paperpile}

\begin{textblock}{7.}(.4,0.95)
\begin{alertblock}{Paperpile}\fontsize{12}{12}\sf
\begin{itemize}\itemsep=2.mm
\item[\textcolor{white}{\ding{229}}] \$3 per month and runs on Google Chrome
\item[\textcolor{white}{\ding{229}}] Papers stored on Google Drive
\item[\textcolor{white}{\ding{229}}] Browser extension for adding papers/books
\item[\textcolor{white}{\ding{229}}] Works with Google Docs or LaTeX.
\item[\textcolor{white}{\ding{229}}] Generate bibliography automatically
\item[\textcolor{white}{\ding{229}}] Handles all formatting for you.
\item[\textcolor{white}{\ding{229}}] Amazingly fast
\end{itemize}
\end{alertblock}
\end{textblock}

\begin{textblock}{7.7}(7.9,6.2)\fontsize{12}{13}\sf
\begin{block}{To install:}
\begin{itemize}
\item Set up account at \url{paperpile.com}
\item Download Google chrome browser extension
\end{itemize}
\end{block}
\end{textblock}\vspace*{10cm}

## What to cite?

\placefig{6.7}{0.}{width=9.35cm}{pileofbooks}\vspace*{0.6cm}

\begin{textblock}{6.5}(0,1.5)
\begin{itemize}\parskip=0cm
\item Cite what is important.
\item Cite (only) what is relevant.
%\item  Cite papers from the journal you intend to submit to.
\item  Avoid lists of gratuitous references.
\item  Include proper citations for all packages and software you use.
\end{itemize}
\end{textblock}\pause

\vspace*{5.2cm}

**When using R**

```r
citation("packagename")
```

## Sight what you cite

\placefig{10.4}{5.4}{width=5.4cm}{bibliography}\vspace*{0.6cm}\vspace*{-0.4cm}
\begin{textblock}{14}(0.2,1.3)
\begin{itemize}\tightlist
\item Every article cited should be sighted, \& preferably read.
 \item  At the very least, check that the article cited really does say what you think it says.
 \item  Type the reference information yourself.
 \item  Don’t just cite what other people say about citations.
 \item  Store accurate reference info from the start.
 \item  Give credit where it is due.

 \only<2>{\vspace*{0.2cm}
   \begin{itemize}\tightlist
   \item Diebold did not invent PITs.
   \item Hyndman did not invent\newline exponential smoothing or\newline ARIMA models.
   \end{itemize}}
\end{itemize}
\end{textblock}

# Searching

## Google Scholar

 * Searching for papers
 * Use advanced search
 * Link GS to your reference manager
 * Track citations of key papers
 * Star papers for your own library
 * Check recommended articles
 * Check author profiles, especially highly cited authors
 * Create your own GS profile once you have (at least) one paper
 * Follow key authors in your area

## Semantic Scholar

\fullwidth{semantic-scholar.png}

# Scripting

## Reproducibility

\begin{textblock}{8.6}(1,1)\fontsize{14}{15}\sf
\begin{block}{Not reproducible:}
\begin{itemize}\tightlist
\item Data edited in a spreadsheet
\item Click and point analysis
\item Copy and paste graphs and tables
\item Tables typed by hand
\end{itemize}
\end{block}
\begin{alertblock}{Reproducible}
\begin{itemize}\tightlist
\item All data edits scripted
\item All analysis scripted
\item Graphs and tables automatically pulled in to the thesis
\item Tables generated with scripts
\end{itemize}
\end{alertblock}
\end{textblock}

\placefig{10.7}{0}{width=5.35cm}{2g9914}

## Reproducibility

\begin{block}{}
Someone should be able to reproduce your thesis without having to guess what software you had installed, what versions, which files do what, etc.
\end{block}\vspace*{-0.2cm}\pause

 * Stay organized.
 * One system for doing this using R is to write your thesis in an Rmarkdown or Quarto file.
 * Track software versions

## Version control
\fontsize{13}{15}\sf

 * `thesis_v1`, `thesis_v2`, etc., is not adequate version control.
 * You need to track changes over time, have a *remote* repository, and be able to roll back as required.
 * Your repository should contain *everything* required to produce your thesis including computer code, references, writing.
 * Your repository should have an obvious structure and be fully documented.
 * **Github** solves these problems
 * Read "Happy git with R": `happygitwithr.com`

\placefig{13}{5.8}{width=1.5cm}{github-mark}
\placefig{11.5}{7.2}{width=4.2cm}{github}

## Version control with git

 * RStudio integrates with github, so version control built in.
 * But github can be used with *any* text-based language including Stata, Python, LaTeX, R, Rmarkdown, Quarto, markdown, etc.
 * Git allows you to:
     - track changes
     - experiment in branches
     - undo
 * Github provides:
    - backup and restore
    - synchronisation

\placefig{13}{5.8}{width=1.5cm}{github-mark}
\placefig{11.5}{7.2}{width=4.2cm}{github}

<!-- ## Some rules for good graphics
\fontsize{14}{16}\sf

1. Design graphics to highlight comparisons you want the reader to make.
1. Choose scales to facilitate comparisons.
1. Use a sensible aspect ratio. width:height of about 1.6, unless axes are on the same scale.
1. Prepare graphics in final aspect ratio. Distorted fonts look awful.
1. Use vector graphics such as eps or pdf.
1. Clear labels and detailed captions.
1. Avoid cluttered legends.
1. Specify units.
1. Label axes.
 -->

# Writing

## Microsoft Word

\placefig{4.1}{1.1}{width=11.9cm}{problems}

\only<2>{
  \begin{textblock}{9}(0.2,1.)
  \begin{block}{}
  \begin{itemize}\tightlist
  \item Impossible to do proper version control
  \item Poor formatting of equations
  \item Impossible to make reproducible
  \item More difficult to automate cross-referencing, numbering
  \item Files more likely to be corrupted
  \end{itemize}\end{block}\end{textblock}}

##

\placefig{1.3}{1.1}{width=3cm}{latex_logo}

\begin{textblock}{7.8}(7.6,.5)\fontsize{13}{13}\sf
\begin{alertblock}{Document processing}
\begin{itemize}\itemsep=1.mm
\item[\textcolor{white}{\ding{229}}] Free and open-source
\item[\textcolor{white}{\ding{229}}] Available on all operating systems
\item[\textcolor{white}{\ding{229}}] Used by every mathematical publisher
\item[\textcolor{white}{\ding{229}}] Separate content from style
\item[\textcolor{white}{\ding{229}}] Format complex equations
\item[\textcolor{white}{\ding{229}}] Automatic numbering
\item[\textcolor{white}{\ding{229}}] Automatic bibliography
\item[\textcolor{white}{\ding{229}}] Almost every language
\end{itemize}
\end{alertblock}
\end{textblock}

\begin{textblock}{5.5}(.7,3.5)
\begin{block}{To install:}
\begin{itemize}
\item Download MikTeX, MacTeX or TeXlive.
\item Download TeXStudio from \url{texstudio.sourceforge.net/}
\end{itemize}
\end{block}
\end{textblock}

## Quarto

\begin{textblock}{9}(6.2,1.5)
\begin{alertblock}{Quarto}
\begin{itemize}\itemsep=1.5mm
\item[\textcolor{white}{\ding{229}}] Combines R, Python  and LaTeX into one system
\item[\textcolor{white}{\ding{229}}] Reproducible research
\item[\textcolor{white}{\ding{229}}] Monash Thesis Template via\\
\texttt{github.com/numbats/monash-quarto-thesis}
\item[\textcolor{white}{\ding{229}}] Useful for assignments too
\item[\textcolor{white}{\ding{229}}] See \url{quarto.org} for help
\end{itemize}
\end{alertblock}
\end{textblock}

\begin{textblock}{5}(.6,1.3)
\begin{block}{}
\includegraphics[width=5cm]{quarto-thesis}
\end{block}
\end{textblock}


# An efficient reproducible workflow

## Example paper

\placefig{0}{1.1}{height=8cm}{figs/jors.png}

\begin{textblock}{9.4}(6.2,1.1)
\begin{block}{}\fontsize{12}{16}\sf
\begin{tabular}{p{5cm}l@{}}
Hyndman RJ, Rostami-Tabar B (2024) Forecasting interrupted time series, \textit{Journal of the Operational Research Society}, in press.
& \raisebox{-3cm}{\includegraphics[width=3.6cm,height=3.6cm]{figs/qr-code}}
\end{tabular}
\end{block}
\end{textblock}
\begin{textblock}{9.4}(6.2,5.4)\fontsize{14}{14}\sf
\begin{alertblock}{}
\href{https://github.com/bahmanrostamitabar/forecasting_interrupted_time_series}{\tt\faicon{github} bahmanrostamitabar/\\\hspace*{1.7ex} \rlap{forecasting\_interrupted\_time\_series}}
\end{alertblock}
\end{textblock}

## Tools

\placefig{4.5}{1.7}{width=1.8cm}{github-mark}
\placefig{3.65}{3.3}{width=3.5cm}{github}
\placefig{8.65}{1.7}{width=3.5cm}{Rlogo}
\placefig{1.15}{4.8}{width=3.5cm}{renv}
\placefig{6.15}{4.8}{width=3.5cm}{targets}
\placefig{11.15}{4.8}{width=3.5cm}{quarto}

## Reproducible environments

\placefig{0.3}{1.25}{height = 6.5cm}{figs/renv}

\begin{textblock}{9.5}(6.2, 1.3)
\begin{itemize}\tightlist
\item Creates project-specific R environments.
\item Uses a package cache so you are not repeatedly installing the same packages in multiple projects.
\item Does not ensure R itself, system dependencies or the OS are the same.
\item Not a replacement for Docker or Apptainer.
\end{itemize}
\end{textblock}

## Reproducible environments
\placefig{0.3}{1.25}{height = 6.5cm}{figs/renv}

\begin{textblock}{9.5}(6.2, 1.3)
\begin{itemize}\tightlist
\item Can use packages from CRAN, Bioconductor, GitHub, Gitlab, Bitbucket, etc.
\item \texttt{renv::init()} to initialize a new project.
\item \texttt{renv::snapshot()} to save state of project to \texttt{renv.lock}.
\item \texttt{renv::restore()} to restore project as saved in \texttt{renv.lock}.
\end{itemize}
\end{textblock}

## Efficient computational pipelines

\placefig{0.3}{1.25}{height = 6.5cm}{figs/targets}

\begin{textblock}{9.5}(6.2, 1.3)
\begin{itemize}\tightlist
\item Supports a clean, modular, function-oriented programming style.
\item Learns how your pipeline fits together.
\item Runs only the necessary computation.
\item Abstracts files as R objects.
\item Similar to Makefiles, but with R functions.
\end{itemize}
\end{textblock}

## Interconnected tasks

\only<1>{\placefig{0.5}{2}{width=13cm}{figs/workflow.png}}
\only<2>{\placefig{0.5}{2}{width=13cm}{figs/change.png}}
\only<3>{\placefig{0.5}{2}{width=13cm}{figs/downstream.png}}
\only<4>{\placefig{0.5}{1.1}{width=14cm}{figs/decisions.png}}

\begin{textblock}{15}(0.1,8.7)
\textcolor{gray}{\fontsize{7}{7}\sf Image from https://wlandau.github.io/targets-tutorial}
\end{textblock}

## Let a pipeline tool do the work

\fullwidth{figs/pipeline_graph.png}\vspace*{-0.15cm}

* Save time while ensuring computational reproducibility.
* Automatically skip tasks that are already up to date.

## Typical project structure

### no_targets.R
\vspace*{-0.26cm}

```{r}
#| eval: false
library(tidyverse)
library(fable)
source("R/functions.R")
my_data <- read_csv("data/my_data.csv")
my_model <- model_function(my_data)
```

\pause

### _targets.R
\vspace*{-0.26cm}

```{r}
#| eval: false
library(targets)
tar_option_set(packages = c("tidyverse", "fable"))
tar_source() # source all files in R folder
list(
  tar_target(my_file, "data/my_data.csv", format = "file"),
  tar_target(my_data, read_csv(my_file)),
  tar_target(my_model, model_function(my_data))
)
```

## Useful targets commands
\vspace*{-0.3cm}

* `tar_option_set()` to set options.
* `tar_target()` to create a target.
* `tar_source()` to source all files in a folder.
* `tar_make()` to run the pipeline.
* `tar_read(object)` to read a target.
* `tar_load(object)` to load a target.
* `tar_visnetwork()` to visualize the pipeline.

\vspace*{0.2cm}\pause

### Random numbers
\vspace*{-0.15cm}

Each target runs with its own seed based on its name and the global seed from `tar_option_set(seed = ???)`


## Reproducible documents

\placefig{0.3}{1.25}{height = 6.5cm}{figs/quarto}

\begin{textblock}{9.5}(6.2, 1.1)
\fontsize{14}{15}\sf
\begin{itemize}\tightlist
\item Generalization of Rmarkdown (not dependent on R)
\item Supports R, Python, Javascript and Julia chunks by using either `knitr`, `jupyter` or `ObservableJS` engines.
\item Consistent yaml header and chunk options.
\item Many output formats, and many options for customizing format.
\item Uses pandoc templates for extensions
\end{itemize}
\end{textblock}

\placefig{5}{7.3}{width=10cm}{figs/qmd.png}

## Chunk options

### Chunk with regular R code

````{verbatim}
```{r}
#| label: fig-chunklabel
#| fig-caption: My figure
mtcars |>
  ggplot(aes(x = mpg, y = wt)) +
  geom_point()
```
````

\pause

### Chunk with targets

````{verbatim}
```{r}
#| label: fig-chunklabel
#| fig-caption: My figure
tar_read(my_plot)
```
````

\only<3>{\begin{textblock}{6}(9,.1)
\begin{alertblock}{}
Reference the figure using \texttt{@fig-chunklabel}.
\end{alertblock}
\end{textblock}}

## targets with quarto

```{r}
#| eval: false
library(targets)
library(tarchetypes)
tar_option_set(packages = c("tidyverse", "fable"))
tar_source() # source all files in R folder
list(
  tar_target(my_file, "data/my_data.csv", format = "file"),
  tar_target(my_data, read_csv(my_file)),
  tar_target(my_model, model_function(my_data)),
  tar_quarto(report, "file.qmd", extra_files = "references.bib")
)
```

1. Load `tarchetypes` package for quarto support.
2. Add a quarto target.

## Extensions and templates
\fontsize{13}{17}\sf

* **Quarto extensions** modify and extend functionality.
  * See <https://quarto.org/docs/extensions/> for a list.
  * They are stored locally, in the `_extensions` folder alongside the qmd document.
* **Quarto templates** are extensions used to define new output formats.
  * **Journal templates** at\newline <https://quarto.org/docs/extensions/listing-journals.html>
  * **Monash templates** at <https://github.com/quarto-monash>

## Slides and links
\fontsize{24}{30}\sf

\vspace*{3cm}

\begin{alertblock}{}
\centerline{\href{https://robjhyndman.com/research_tools}{\tt\bfseries robjhyndman.com/research\_tools}}
\end{alertblock}
